Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-05-26 FPT2014/bare_conf.tex
Feedbacks from George

\begin{abstract}
%\boldmath
Online arithmetic has been widely utilized for ASIC implementations, as it is designed to perform computations from the most significant digits. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to operators with conventional arithmetic. Unfortunately, online arithmetic operators require a large area overhead, while efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement key primitives of online arithmetic, such as adders and multipliers, efficiently on modern FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that the proposed architectures achieve significant area savings and frequency speed-ups compared to the original implementations.

@@

Recent research has demonstrated another key advantage of online arithmetic is that it allows gracefully degradation when operating beyond the deterministic region, e.g. under overclocking \cite{SKDAC14}. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional forms of computer arithmetic do not fail gracefully when timing violation happens, because timing errors initially affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occur at LSDs of the results.

Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to the huge area overhead. Generally, it is difficult to efficiently map the arithmetic circuits solely onto the look-up-tables (LUTs), which are fundamental building blocks of FPGAs. Lots of efforts has been done to alleviate this concern. For instance, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. There is a large amount of studies exploring methods that can effectively map circuits using the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or propose alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2}. Furthermore, hard DSP blocks are also included to boost multiplications and multiply-accumulations~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, we notice from existing literature that there are two directions worth investigating. First, although there has been previous research about FPGA implementations of online arithmetic, they typically only focus on online adders. A brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration. Second, we found that existing approaches of building general purpose online adders are specifically for low cost FPGAs with 4-input LUTs and 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan~3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and 4 LUTs in a Slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a Slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. In addition, the theoretical resource usage of both operators is given. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. In summary, the main contributions of this paper are as follows:
\begin{enumerate}
    \item An efficient approach of mapping online adders onto modern FPGAs;
    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
    \item Theoretical area models of the proposed architectures.
\end{enumerate}

%The theoretical area overhead of an online adder is given in terms of the usage of LUTs and Slices.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    \vspace{-2ex}
\end{figure}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA Slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of the \texttt{RCA} drops drastically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Furthermore, our approach operates faster than \texttt{OA\_behv}; the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $20\%$ to $26\%$.

@@

@@

\section{Digit Parallel Online Multiplier on FPGAs}\label{Sec:OM_FPGA}
\subsection{Algorithm of Digit Parallel Online Multiplication}
Algorithm \ref{Algorithm:OnlineMult} as described in Section~\ref{Sec:Background} can be synthesized into a unrolled digit parallel structure. Recent study has shown that the digit-parallel online multiplier (\texttt{OM}) is also more tolerant to timing violations in comparison to the conventional arithmetic, because timing errors initially affect the LSDs and this will lead to smaller error magnitude~\cite{SKDAC14}. This implementation method is straightforward and can be directly utilized to design other digit parallel online operators.

On the other hand, significant area overhead is required in order to implement this type of \texttt{OM}. This is due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, or $(N+\delta)$ iterations are required to generate the most significant $N$ digits. Second, the logic of each stage is complex as it involves operations such as online addition, digit-vector multiplication, shifting and other combination logic blocks such as the selection function. Also, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $(N+2+\delta)$, because the selection function needs 2 integer digits and 1 fractional digits to generate $z_j$, as described in Section~\ref{Sec:Background}. In comparison, an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.

We address the first 2 design issues by providing an alternative online multiplication algorithm. Instead of simply implementing the unrolled format of Algorithm \ref{Algorithm:OnlineMult}, it can be optimized specifically for digit parallel operations. We notice that the online delay $\delta$ is employed in Algorithm~\ref{Algorithm:OnlineMult} because the input data is available in a digit serial fashion. Hence $\delta$ is used to accumulate enough input digits to generate the MSD of the result. Nevertheless if all digits of inputs are available simultaneously, $\delta$ is no longer necessary. Due to the same reason, $H[j]$, which takes one digit of each input per stage as shown in the original algorithm, can be optimized to take one partial product $Xy_j$ or $Yx_j$ per stage. To summarize, the modified algorithm for digit parallel online multiplication is described in Algorithm~\ref{Algorithm:OM_DigitParallel}, where $N$ denotes the operand word-length, function $frac(W_{[N]})$ refers to select the fractional digits of $W_{[N]}$, and the selection function $sel()$ is given previously in (\ref{Eq:SelFunc_OM}).

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $Z_{[j]}  ~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$
  \end{algorithmic}
\end{algorithm}

In comparison to Algorithm~\ref{Algorithm:OnlineMult}, the maximum stages required to generate results with full precision drops by a factor of 2 using the proposed new algorithm. In each stage, the word-length of all signals also reduces from $(N+2+\delta)$ to $(N+2)$, as $\delta$ is not necessary for digit parallel operations. In addition, the $frac()$ function can be implemented only based on wire connections without using other logic resources.

@@

The theoretical minimum resource uses of the proposed \texttt{OM} can be calculated as follows. In an $N$-digit \texttt{OM}, overall $(N-1)$ online adders are needed. According to Algorithm~\ref{Algorithm:OM_DigitParallel}, the word-length of each online adder is $(N+2)$-digit (2 integer digits and $N$ fractional digits). Therefore based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), the number of LUTs and the number of Slices used by the \texttt{OM} is given in (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}), respectively. Note that $L_{S1}$ and $S_{S1}$ denote the number of LUTs and Slices used by the first stage, which is built without online adders.
%
\begin{eqnarray}\label{Eq:OMareaLUT}
	OM\_{LUT} = 2(N+2)(N-1) + {L}_{S1}
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OMareaSlice}
	OM\_{Slice} = (1+\left\lceil{\frac{N}{2}}\right\rceil)(N-1)+S_{S1}
\end{eqnarray}

@@

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/ParallelMult_MSDhalf.pdf}
	\caption{(a) Modified structure of a 4-digit online multiplier which only generates the most significant 4-digit result. (b) Structure diagram of stage $j$, with the word-lengths of all internal signals labelled.}
	\label{Fig:PMStructure_MSDhalf}
	\vspace{-2ex}
\end{figure}

\subsection{Performance Analysis}
Similar to the experiments as described in Section~\ref{subsec:OA_Performance}, we compare the online multiplier based on behavioral descriptions (\texttt{OM\_initial}) against the proposed 2 types of multipliers, which are implemented to generate products with full precision (\texttt{OM\_full}) and half precision (\texttt{OM\_half}), respectively. In addition, we also compare the \texttt{OM} with the conventional binary arithmetic multiplier. The conventional multiplier is created using the Xilinx Core Generator with speed optimization~\cite{XilinxMult}, and it is implemented based on LUTs without DSPs.

@@

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_LUTslog2.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_Slicelog2.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different types of binary multipliers. (a) Usage of LUTs. (b) Usage of Slices}
  \label{Fig:OMArea}
  \vspace{-2ex}
\end{figure}

We also check the rated frequencies of \texttt{OM\_full}, \texttt{OM\_half} and \texttt{CoreGen} for a variety of operand word-lengths. The frequency values are still obtained from the post place and route timing reports. Notice that input stimulus to the circuits are randomly sampled from a uniform distribution of $N$-digit numbers. The results are plotted in Fig.\ref{Fig:OM_Freq}, from which we can see that there is an increasing gap between the online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This might because in the online multiplier, although the carry resources in the FPGA are efficiently used for area reduction, each 4-digit carry logic within a Slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is huge for large operand word-lengths. This also explains the fact that \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, as the critical path of both designs are identical in theory.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.48\textwidth]{./Figures/Exp/OM_Freq.eps}
	\caption{Rated frequencies of different types of binary multipliers for a variety of operand word-lengths. The results are obtained from the Xilinx Timing Analyzer in ISE 14.1 after placing and routing.}
	\label{Fig:OM_Freq}

@@

However, we can take advantage of the ``overclocking friendly'' feature of the online arithmetic, and push the previous tests one step further by operating the circuits beyond their deterministic region while allowing timing violations to happen. We examine the overclocking behavior of all 3 types of multipliers in terms of the mean relative error (\texttt{MRE}), which is given in (\ref{Eq:MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of correct output, respectively.
%
\begin{eqnarray}\label{Eq:MRE}
  MRE=\frac{E_{error}}{E_{out}}\times100\%
  \vspace{-1ex}
\end{eqnarray}

For instance, Fig.~\ref{Fig:PM_MRE} demonstrates the \texttt{MRE} values of the 8-digit multipliers. We see that although the Timing Analyzer predicts slower frequencies for the \texttt{OM}s, they actually operate faster than \texttt{CoreGen} without the occurrence of timing errors. This is in accordance with the results seen in the previous work \cite{SKDAC14}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.48\textwidth]{./Figures/Exp/MRE_PM.eps}
 % \vspace{-1ex}
  \caption{Mean relative errors seen at the outputs of different types of binary multipliers, when clocked with faster-than-rated frequencies.}
  \label{Fig:PM_MRE}
  \vspace{-1ex}
\end{figure}

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementations. It is designed to perform computations from the most significant digits, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement key primitives of online arithmetic, such as adders and multipliers, efficiently on modern FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ Slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ area savings and $1.5\times$ speed-ups for multipliers.

% significant area savings and frequency speed-ups compared to a direction.

@@

Recent research has demonstrated another key advantage of online arithmetic: that it allows graceful degradation when operating beyond the deterministic clocking region, i.e. under overclocking \cite{SKDAC14_REVIEW}. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional forms of computer arithmetic do not fail gracefully when timing violation happens, because timing errors tend to affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occur at LSDs of the results.

Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to a large area overhead. For standard arithmetic, it is difficult to efficiently map arithmetic circuits solely onto look-up-tables (LUTs), the fundamental building blocks of FPGAs. As a result, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. Significant work explores methods that can effectively map circuits to the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or which proposes alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2}. Furthermore, hard DSP blocks are also included in modern architectures to improve the area and performance of multiplications and multiply-accumulates~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, two research directions have been investigated. First, although there has been previous research about FPGA implementations of online arithmetic, the work only focused on online adders; a brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. Second, there are existing approaches to build general purpose online adders are specifically for low cost FPGAs with 4-input LUTs and 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan~3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and 4 LUTs in a Slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}. Moreover, the move to the 6-LUT on the basic architecture opens new opportunities for efficient implementation.
%However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a Slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. In addition, the resource usage of both operators is given. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ Slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. For online multipliers, the LUTs and Slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. In summary, the main contributions of this paper are an efficient mapping of online adders and multipliers to modern FPGAs. (to be completed)
% by fully utilizing the fast carry logic for .

%\begin{enumerate}
%    \item An efficient approach of mapping online adders onto modern FPGAs;
%    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
%    \item Theoretical area models of the proposed architectures.
%\end{enumerate}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    %\vspace{-2ex}
\end{figure}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA Slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	%\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

@@

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of the \texttt{RCA} drops drastically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Furthermore, our approach operates faster than \texttt{OA\_behv}; the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $1.20\times$ to $1.26\times$.

@@

\section{Digit Parallel Online Multiplier on FPGAs}\label{Sec:OM_FPGA}
\subsection{Algorithm of Digit Parallel Online Multiplication}
Algorithm \ref{Algorithm:OnlineMult} as described in Section~\ref{Sec:Background} can be synthesized into a unrolled digit parallel structure. A recent study has shown that the digit-parallel online multiplier (\texttt{OM}) is also more tolerant to timing violations in comparison to the conventional arithmetic, because timing errors initially affect the LSDs and this will lead to smaller error magnitude~\cite{SKDAC14_REVIEW}. %This implementation method is straightforward and can be directly utilized to design other digit parallel online operators.

However, significant area overhead is required in order to implement this type of \texttt{OM}, due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, or $(N+\delta)$ iterations are required to generate the most significant $N$ digits. Second, the logic of each stage is complex as it involves operations such as online addition, digit-vector multiplication (to generate $x_{j+\delta+1}\cdot Y_{[j+1]}$ and $y_{j+\delta+1}\cdot X_{[j]}$ in Algorithm~\ref{Algorithm:OnlineMult}), shifting and other combinational logic blocks such as the selection function. Also, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $(N+2+\delta)$, because the selection function needs 2 integer digits and 1 fractional digit to generate $z_j$, as described in Section~\ref{Sec:Background}. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.
%In comparison, an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder.

We address the first two design issues by providing an alternative online multiplication algorithm. Instead of simply implementing the unrolled format of Algorithm \ref{Algorithm:OnlineMult}, it can be optimized specifically for digit parallel operations. We notice that the online delay $\delta$ is employed in Algorithm~\ref{Algorithm:OnlineMult} because the input data is available in a digit serial fashion. Hence $\delta$ is used to accumulate enough input digits to generate the MSD of the result. However if all digits of the inputs are available simultaneously, then the delay $\delta$ is no longer necessary. For the same reason, $H[j]$, which takes one digit of each input per stage as shown in the original algorithm, can be optimized to take one partial product $Xy_j$ or $Yx_j$ per stage. To summarize, the modified algorithm for digit parallel online multiplication is described in Algorithm~\ref{Algorithm:OM_DigitParallel}, where $N$ denotes the operand word-length, function $frac(W_{[N]})$ refers to select the fractional digits of $W_{[N]}$, and the selection function $sel()$ is given previously in (\ref{Eq:SelFunc_OM}).

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $z_{j}  ~~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$
  \end{algorithmic}
\end{algorithm}

In comparison to Algorithm~\ref{Algorithm:OnlineMult}, the maximum number of stages required to generate results with full precision drops by a factor of 2 using the proposed new algorithm. In each stage, the word-length of all signals also reduces from $(N+2+\delta)$ to $(N+2)$, as $\delta$ is not necessary for digit parallel operations. In addition, the $frac()$ function can be implemented only based on wire connections without using other logic resources.

@@

The theoretical minimum resource uses of the proposed \texttt{OM} can be calculated as follows. In an $N$-digit \texttt{OM}, overall $(N-1)$ online adders are needed. According to Algorithm~\ref{Algorithm:OM_DigitParallel}, the word-length of each online adder is $(N+2)$-digit (2 integer digits and $N$ fractional digits). Therefore based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), the number of LUTs and the number of Slices used by the \texttt{OM} is given in (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}), respectively. Note that $L_{S1}$ and $S_{S1}$ denote the number of LUTs and Slices used by the first stage, which is built without online adders.

\begin{eqnarray}\label{Eq:OMareaLUT}
	OM\_{LUT} = 2(N+2)(N-1) + {L}_{S1}
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OMareaSlice}
	OM\_{Slice} = (1+\left\lceil{\frac{N}{2}}\right\rceil)(N-1)+S_{S1}
\end{eqnarray}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/ParallelMult_MSDhalf.pdf}
	\caption{(a) Modified structure of a 4-digit online multiplier which only generates the most significant 4-digit result. (b) Structure diagram of stage $j$, with the word-lengths of all internal signals labelled.}
	\label{Fig:PMStructure_MSDhalf}
	%\vspace{-2ex}
\end{figure}

@@

\subsection{Performance Analysis}
Similar to the experiments as described in Section~\ref{subsec:OA_Performance}, we compare the online multiplier based on behavioral descriptions (\texttt{OM\_behv}) against the proposed 2 types of multipliers, which are implemented to generate products with full precision (\texttt{OM\_full}) and half precision (\texttt{OM\_half}), respectively. In addition, we also compare the \texttt{OM} with the conventional binary arithmetic multiplier. The conventional multiplier is created using the Xilinx Core Generator with speed optimization~\cite{XilinxMult}, and it is implemented based on LUTs without DSPs.

@@

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_LUTslog.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_Slicelog.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different types of binary multipliers. (a) Usage of LUTs. (b) Usage of Slices}
  \label{Fig:OMArea}
  %\vspace{-2ex}
\end{figure}

We also check the rated frequencies of all four types of multipliers for a variety of operand word-lengths. The frequency values are still obtained from the post place and route timing reports. Notice that input stimulus to the circuits are randomly sampled from a uniform distribution of $N$-digit numbers. The results are plotted in Fig.\ref{Fig:OM_Freq}, from which we can see that in comparison to \texttt{OM\_behv}, our proposed architectures achieve frequency speed-ups varying from $1.49\times$ to $1.98\times$.

We also notice that there is an increasing gap between the proposed online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This might because in the online multiplier, although the carry resources in the FPGA are efficiently used for area reduction, each 4-digit carry logic within a Slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is huge for large operand word-lengths. This also explains the fact that \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, as the critical path of both designs are identical in theory.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.48\textwidth]{./Figures/Exp/OM_Freq2.eps}
	\caption{Rated frequencies of different types of binary multipliers for a variety of operand word-lengths. The results are obtained from the Xilinx Timing Analyzer in ISE 14.1 after placing and routing.}
	\label{Fig:OM_Freq}

@@

However, we can take advantage of the ``overclocking friendly'' feature of the online arithmetic, and push the previous tests one step further by operating the circuits beyond their deterministic region while allowing timing violations to happen. We examine the overclocking behavior of all 3 types of multipliers in terms of the mean relative error (\texttt{MRE}), which is given in (\ref{Eq:MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of correct output, respectively.

\begin{eqnarray}\label{Eq:MRE}
  MRE=\frac{E_{error}}{E_{out}}\times100\%
  %\vspace{-1ex}
\end{eqnarray}

For instance, Fig.~\ref{Fig:PM_MRE} demonstrates the \texttt{MRE} values of the 8-digit multipliers. We see that although the Timing Analyzer predicts slower frequencies for the \texttt{OM}s, they actually operate faster than \texttt{CoreGen} without the occurrence of timing errors. This is in accordance with the results seen in the previous work \cite{SKDAC14_REVIEW}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.48\textwidth]{./Figures/Exp/MRE_PM.eps}
 % \vspace{-1ex}
  \caption{Mean relative errors seen at the outputs of different types of binary multipliers, when clocked with faster-than-rated frequencies.}
  \label{Fig:PM_MRE}
  %\vspace{-1ex}
\end{figure}
